---
title: "Assignment 3: Data Exploration"
author: "Yanxi Peng"
output: pdf_document
geometry: margin=2.54cm
---

## OVERVIEW

This exercise accompanies the lessons in Environmental Data Analytics on Data Exploration.

## Directions

1.  Rename this file `<FirstLast>_A03_DataExploration.Rmd` (replacing `<FirstLast>` with your first and last name).
2.  Change "Student Name" on line 3 (above) with your name.
3.  Work through the steps, **creating code and output** that fulfill each instruction.
4.  Be sure to **answer the questions** in this assignment document.
5.  When you have completed the assignment, **Knit** the text and code into a single PDF file.
6.  After Knitting, submit the completed exercise (PDF file) to the dropbox in Sakai.

The completed exercise is due on Sept 30th.

## Set up your R session

1.  Check your working directory, load necessary packages (tidyverse), and upload two datasets: the ECOTOX neonicotinoid dataset (ECOTOX_Neonicotinoids_Insects_raw.csv) and the Niwot Ridge NEON dataset for litter and woody debris (NEON_NIWO_Litter_massdata_2018-08_raw.csv). Name these datasets "Neonics" and "Litter", respectively. Be sure to include the subcommand to read strings in as factors.

```{r,  message = FALSE}
library(tidyverse)
getwd()
setwd('C:/Users/16920/Documents/R/EDA-Fall2022')
Neonics <- read.csv("./Data/Raw/ECOTOX_Neonicotinoids_Insects_raw.csv", stringsAsFactors = TRUE)
Litter <- read.csv("./Data/Raw/NEON_NIWO_Litter_massdata_2018-08_raw.csv", stringsAsFactors = TRUE)


```

## Learn about your system

2.  The neonicotinoid dataset was collected from the Environmental Protection Agency's ECOTOX Knowledgebase, a database for ecotoxicology research. Neonicotinoids are a class of insecticides used widely in agriculture. The dataset that has been pulled includes all studies published on insects. Why might we be interested in the ecotoxicology of neonicotinoids on insects? Feel free to do a brief internet search if you feel you need more background information.

> Answer: Neonicotinoids may harm other insects other than pests or even cause widespread death of certain types of insects.

3.  The Niwot Ridge litter and woody debris dataset was collected from the National Ecological Observatory Network, which collectively includes 81 aquatic and terrestrial sites across 20 ecoclimatic domains. 32 of these sites sample forest litter and woody debris, and we will focus on the Niwot Ridge long-term ecological research (LTER) station in Colorado. Why might we be interested in studying litter and woody debris that falls to the ground in forests? Feel free to do a brief internet search if you feel you need more background information.

> Answer: Because litter and woody debris plays an important role in carbon budgets and nutrient cycle which will effect the water flows process and ecosystem.

4.  How is litter and woody debris sampled as part of the NEON network? Read the NEON_Litterfall_UserGuide.pdf document to learn more. List three pieces of salient information about the sampling methods here:

> Answer: 
 1.Litter and fine woody debris sampling is executed at terrestrial NEON sites that contain woody vegetation >2m
tall 
 2.Sampling for this product occurs only in tower
plots
 3. Locations of tower plots are selected randomly within the 90% flux footprint of the primary and secondary airsheds

## Obtain basic summaries of your data (Neonics)

5.  What are the dimensions of the dataset?

```{r}
dim(Neonics)
```

6.  Using the `summary` function on the "Effect" column, determine the most common effects that are studied. Why might these effects specifically be of interest?

```{r}
summary(Neonics$Effect)
```

> Answer: The most common effects that are studied is Population. Because different type of Neonicotinoids can have different effects on bees, these effects could lead to the extinction of specific species of bees or have an ecological impact on their survival.

7.  Using the `summary` function, determine the six most commonly studied species in the dataset (common name). What do these species have in common, and why might they be of interest over other insects? Feel free to do a brief internet search for more information if needed.

```{r}
summary(Neonics$Species.Common.Name)
```

> Answer: The six most commonly studied species in the dataset in Honey Bee, Parasitic Wasp, Buff Tailed Bumblebee. Carniolan Honey Bee, Bumble Bee and Italian Honeybee. Because they are systemic chemicals absorbed into the plant, neonicotinoids can be present in pollen and nectar, making them toxic to pollinators that feed on them.  REF:http://www.xerces.org/publications/scientific-reports/how-neonicotinoids-can-kill-bees

8.  Concentrations are always a numeric value. What is the class of Conc.1..Author. in the dataset, and why is it not numeric?

```{r}
class(Neonics$Conc.1..Author.)
```

> Answer: Factor. Because it contain strings.

## Explore your data graphically (Neonics)

9.  Using `geom_freqpoly`, generate a plot of the number of studies conducted by publication year.

```{r}
ggplot(Neonics) +
  geom_freqpoly(aes(x=Publication.Year), bins=50)
```

10. Reproduce the same graph but now add a color aesthetic so that different Test.Location are displayed as different colors.

```{r}
ggplot(Neonics) +
  geom_freqpoly(aes(x=Publication.Year,color = Test.Location), bins=50)
```

Interpret this graph. What are the most common test locations, and do they differ over time?

> Answer: The most text location is lab, and as the publication year increase the test location at lab increase. And the highest peak was observed around Year 2015.

11. Create a bar graph of Endpoint counts. What are the two most common end points, and how are they defined? Consult the ECOTOX_CodeAppendix for more information.

```{r}
ggplot(Neonics, aes(x = Endpoint)) +
  geom_bar()
```

> Answer: The two most common end points are NOEL and LOEL. NOEL:No-observable-effect-level: highest dose (concentration) 
producing effects not significantly different from responses of 
controls according to author's reported statistical test
LOEL:Lowest-observable-effect-level: lowest dose (concentration)
producing effects that were significantly different (as reported
by authors) from responses of controls.

## Explore your data (Litter)

12. Determine the class of collectDate. Is it a date? If not, change to a date and confirm the new class of the variable. Using the `unique` function, determine which dates litter was sampled in August 2018.

```{r}
library("lubridate")
class(Litter$collectDate)
Litter$collectDate <- ymd(Litter$collectDate)
class(Litter$collectDate)

unique(Litter$collectDate)
```

13. Using the `unique` function, determine how many plots were sampled at Niwot Ridge. How is the information obtained from `unique` different from that obtained from `summary`?

```{r}
unique(Litter$plotID)

summary(Litter$plotID)
```

> Answer: 12 Levels. 'summary' function gives the total number included in each plots, 'unique' function gives the total number of different type of plots.

14. Create a bar graph of functionalGroup counts. This shows you what type of litter is collected at the Niwot Ridge sites. Notice that litter types are fairly equally distributed across the Niwot Ridge sites.

```{r}
ggplot(Litter, aes(x = functionalGroup)) +
  geom_bar(fill = 'firebrick1')
```

15. Using `geom_boxplot` and `geom_violin`, create a boxplot and a violin plot of dryMass by functionalGroup.

```{r}
#boxplot
ggplot(Litter) +
  geom_boxplot(aes(x =functionalGroup , y = dryMass))
#violin plot
ggplot(Litter) +
  geom_violin(aes(x = functionalGroup, y = dryMass))
```

Why is the boxplot a more effective visualization option than the violin plot in this case?

> Answer: Beacuse in this dataset, the data is scattered, causing the violin plot to have an abnormal shape.

What type(s) of litter tend to have the highest biomass at these sites?

> Answer: Needles
